Improving content discovery through combining linked data and data mining techniques

### 2016-05-02 report/chapters/06-evaluation.tex
First pass of chapter 6 (minus survey results)

\section{Qualitative Survey}

\subsection{Survey Design}
\subsection{Survey Results}
\subsection{Further Observations and Discussion}

@@

\section{Implementation Difficulties}

>>>>

This chapter builds on the more immediate analysis in
chapter~\ref{chp:analysis} of the clusters produced by evaluating the
results across a group of people. The system designed and built is
also a subject of evaluation, particularly any flaws or potential
improvements that emerge from having run the software and reviewed its
results.

A higher-level
goal of this research was to perform some more qualitative evaluation
of those clusters by obtaining human feedback on a potential
application thereof. The application evaluted was a feature that would
take a page a user is currently reading and suggest related content
based on their being members of the same cluster. Mock-ups of such
a feature were generated and evaluated via a survery as described
in section~\ref{sec:survery}.

The remaining sections of this chapter then discuss some issues with
the design and difficulties arising during implementation
respectively.

\section{Qualitative Survey}
\label{sec:survey}

An online survey was produced where a user is shown a sample page
from the BBC website and for each cluster of which that page is
a member, a row of up to four ``suggestions'' was presented to users.

In this section, some details around the design and sampling criteria
for the survey are discussed initially, followed by analysis of the
results. The section concludes with further qualitative feedback and
observations that arose in survey responses.

\subsection{Survey Design}

An initial challenge for generating the survey was that it was
necessary to present to users a view starting from a particular
content item. This as effectively an inversion of the analysis
from chapter~\ref{chp:analysis} where a top-down overview of the
clusters was performed, with some drill-down into more depth in some
cases.

A reasonably arbitrary sample size of 20 content items was chosen
for which it was therefore necessary to choose 20 \emph{pages} that
were known to be members of one or more clusters. It was also
desirable to ensure that multiple clusters are evaluated in one
survey question, so items that were placed into clusters by multiple
instances of the data pipeline were preferable.

It was also important to ensure that the overall set of items
chosen provided coverage across all permutations of the data pipeline
and that some diversity of content was present (e.g. all items are
not simply news articles without any representation of TV and video
content).

A further additional property chosen for the sample set was that
a diverse range of cluster \emph{cohesion} values were present in the
set. This arguably allows the survey to present for evaluation
clusters that the machine learning applied itself believes to be poor
as well as ones it believes to be cohesive. This controls
for coincidences where only good clusters were chosen for one approach
and another had its least cohesive results chosen. It also has
potential not just to compare which approaches receive the most
relevance votes, but also to look for where those approaches' results
produce clusters whose cohesion measure correlate with anything
meaningful to humans.

The above sampling heuristics were applied programmatically by first
inverting the cluster data structures such that there is a known
list of all content IRIs across the whole results set and the list
of clusters of which they are members are keyed against those IRIs.

A composite uniqueness key was then created for each membership those
content IRIs have in the form:

$$
u(iri, cluster) = (approach, cohesion(cluster), category(iri))
$$

\noindent where $approach$ is a unique key referring to a particular
permutation of the data pipeline being evaluated and $category$ is
a rough, heuristic function that tries to assess which part of the BBC
website ``owns'' that content item, e.g. BBC News, iPlayer, Weather,
Sport. Each item would have a uniqueness key attached to it for each
cluster of which it is a member.

Content items were initially scored based on how many clusters of
which they were a member:

$$
score(iri) = 10 - |5 - \left\vert{\{c \in C \: | \: iri \in c\}}\right\vert|
$$

\noindent where $C$ is the set of all clusters produced. This gives
us a score of 10 when the number of clusters in which the item appears
is the ideal value of five. The justification for this optimum is
based on the lower end of the memory capacity limits suggested by
Miller\cite{miller1956magical} (Miller's Law).

The item is thus ``punished'' when it appears in only one cluster
(thus the question would only be testing one approach and has less
value as a question) or if the item appears in too many clusters
(asking people to compare between too many items might be
overwhelming and reduce the potential for people to answer
accurately). As this is only a heuristic, the sampling code still
chose some items nearer this extremes, thus an absolute cap was
implemented in the survey itself to ensure a random sample of seven
clusters was chosen for evaluation when items appeared in a number
greater than seven.

Starting with the highest scoring so far, the sampling algorithm
then updated each score based on the number of unseen unique keys
it would introduce to the final sample if it were chosen:

$$
score'(iri) = score(iri) \times \frac{15 - |keys(iri) \cap K|}{2}
$$

\noindent where $keys(iri)$ is all uniqueness keys for an IRI:

$$
keys(iri) = \bigcup_{c \in C} u(iri, c)
$$

\noindent and $K$ is the cumulative union of all keys so far
generated by invocations of of the $keys$ function so far, with its
initial value being the empty set.

Some of the constants used were tweaked based on what appeared to
generated a desirable, mixed set of sample results such that they are
a little better than arbitrary values. Once all IRIs were updated with
the new scores, the final sample of
results for evaluation was simply the content items with the twenty
highest scores.

\subsection{Survey Results}

TODO

\subsection{Further Observations and Discussion}

TODO

@@

Throughout chapter~\ref{chp:design}, the design of the data pipeline
and overall software architecture was outlined and known trade-offs
and limitations were highlighted. Further design trade-offs due to
the systems intended use to be experimental rather than
production-quality in an enterprise were also covered.

This section highlighted some design issues that were noted only after
the system was built and how research and industry alike might learn
from these problems. Such problems might be addressed if the results
of this experiment are to be researched further or indeed if such
a system were to be built in a real media organisation. Such
learnings for production use should follow on from the
recommendations in section~\ref{sec:production-recommendations} around
applying the data extraction techniques used and provide more general
software design recommendations.

\subsection{Performance}

The experimental system ran far slower than would be acceptable in
an enterprise system and even to the extent that it would limit
developing this research to the scale where it is processing millions
of items -- a target that would be ideal for further proving the
suitability of semantics and machine learning on real media
content corpora.

A better-performing system might have allowed even this experiment
to consider larger numbers of items, although even a relatively
small sample size should sufficiently indicate which techniques look
more promising for developing a more efficient system that focuses
thereon (and save time by dismissing techniques that show no promise).

The performance difficulties that arose from the design and concept
broadly break down into two problems:

\begin{enumerate}
\item lack of consideration for data structures for the feature sets
  and clusters; and
\item network latency bottlenecks.
\end{enumerate}

The first problem stems from the fact that the design was precise
around the in-flight data in the pipeline, less consideration was
given for how to store the resulting feature sets and also the
clusters. The feature sets were stored as simple JSON objects on
files on disk, which is perfectly fast for writing and reading, but
there was no pre-calculated distance matrix between those objects. The
clusters were also stored in a na\"ive fashion of larger JSON objects
that embedded their members' data for quick display of those clusters
and calculation of things like cohesion.

A distance matrix is a $O(n^2)$ operation in terms of computational
complexity, which scales poorly, particularly for a continuous ingest
of data where each new item invalidates the matrix. The design of
the system implemented in this research was built around an assumption
that we wish to have a continuously-running data pipeline that can
update as new content is published. Recalculating the distance matrix
on each new item would effectively push the complexity up to $O(n^3)$
which is clearly undesirable.

With this in mind, a cut-off for content ingest had to be determined
so that the distance calculations and then the clustering could be
invoked. The system implemented does not save any intermediate data
such as distances, so adding in a single new item or cleaning out
any duplicates found required all the results to be regenerated from
the start.

A production version of this system would have to look at a distance
matrix and cluster set structure that allows for real time updates
of new and updated content, efficiently changing only parts of the
matrix affected (e.g. a new item only needs to calculate its distances
with each existing item and does not affect existing pairs). This
could be a promising direction for further research.

The other issue of network latency became very apparent in particular
when enrichment via object dereferencing was enabled. Whereas an
enterprise integration effort between two different data stores
operates on bespoke software engineered in a controlled network,
extracting RDF data via a URL dereference is susceptible to network
latency over the public Internet. Also, HTML pages with embedded
semantics have a very poor ratio of semantic metadata to irrelevant
information -- a document DOM focuses a lot on marking up the content
for web browsers.

There is still a compelling case for the \emph{decoupling} achieved
by using standards like RDFa and the \emph{abstraction} HTML and HTTP
provide over the underlying data stores, reducing the need for
bespoke clients. However, any extraction of that data needs to
tolerate or factor in the increased latency for this approach.

For extraction, the latency of a single HTTP request (or two or three
if HTTP redirects are issued) will only cause a bottleneck if new
content is published or updated at a rate faster than the pipeline
can process. This seems unlikely even in an organisation like the BBC,
particularly if the system implements structures such as queues to
handle bursts of input, processing them in quieter intervals.

However, for enrichment via deference, an RDF graph produced might
have hundreds of objects needing dereferenced via HTTP request. A
simple cache layer might prevent the system repeatedly requesting for
the same URL in a short period, but this kind of latency still means
it could take of the order of minutes for a single item to get
through the pipeline.

It is not clear that there is a single answer to handle this as it is
likely a combination of factors need optimised to make this form
of enrichment acceptable:

\begin{itemize}
\item It might be that hundreds of objects linked via predicates
  creates too much noise in general. If those objects have come mainly
  from entity extraction, should entity extraction be tuned to bring
  back only the ``main'' topics of a page?
\item Chapter~\ref{chp:analysis} discussed how many of the extraction
  techniques (particularly deference) produced a lot of irrelevant
  facts (e.g. about the page itself, not the content) all of which
  would be candidiates for deference of the objects are IRIs.
  Effectively filtering out such noise would reduce the work spent
  enriching irrelevant things further.
\item A simple architectural solution is to parallelise the HTTP
  calls as much as possible such that an item's latency is only as
  long as the slowest response time. There are obvious limits to this
  (we should avoid overloading the server even if it is scaled for
  millions of visitors per day), but any parallelism is clearly
  better than a hundred sequential HTTP requests. Parallelism includes
  distributed architecture over multiple servers as much as
  concurrent programming paradigms.
\item An organisation should probably ensure a system is physically
  near any servers serving HTTP requests to reduce the overall TCP
  packet latency. This optimisation will add up over time, but is
  a challenge for a large organisation such as the BBC with systems
  of varying ages distributed over multiple locations
  \footnote{Industry has
  long outgrown the notion
  that enterprises like the BBC have a single ``www'' server as
  in the early days of the World Wide Web. However, the www hostname
  still needs to resolve to a single set of IP addresses for something
  that fronts the rest of the network like a load balancer. Effective
  caching in this front layer and ensuring our data pipline lives close
  to that cache might be sufficient in this case. An example of a
  front cache might be something like the Varnish HTTP cache:
  https://www.varnish-cache.org/}
\end{itemize}

Even with all the optimisations possible, it is unescapable that
multiple HTTP requests to assemble an RDF graph is slower than a
single query to a data store's query engine (e.g. SQL) that returns
all the information needed in a single response. The benefits of using
semantics, HTTP and other standards -- such as decoupling, simple
enrichment, reuse of standard technologies -- need to be effectively
researched and promoted, otherwise enterprise organisations will
continue to reach for the seemingly obvious solution of building
software that queries one database to push to another.

\subsection{Noise}

The semantic web and more recently Linked Data are arguably built
around the idea that they embrace the loose architecture of the
World Wide Web. With that looseness -- open world assumption,
no schema constraints, etc. -- we cannot assure strict quality and
integrity of the data as would be possible in a closed-world
database.

The data pipeline design in chapter~\ref{chp:design} has no
requirements nor consideration for how to handle facts that are
erroneous, missing nor meaningless (e.g. see the discussion
in section~\ref{sec:anal-deref} of the system finding facts about the
pages' structure rather than content metadata).

In section~\ref{sec:impact-of-noise}, it is argued that noise might
not have as much impact as one would assume given the fuzzy nature
of machine learning in general. This argument is important to consider,
but here we explore the issue of noise on the assumption that it
\emph{is} a problem as it is important to evaluate how noise fed into
the system in the first place and how we might tackle it.

Erroneous facts may be non-trivial to distinguish from real facts, but
there is scope for research on effective methods for reducing that
noise. In an enterprise, we need not exclude the benefits of
enterprise integration development to \emph{support} this approach
based on semantics and web standards.

Whilst there is much gained
from being able to reason about the entire breadth of an organisation's
content using semantics, it does not preclude writing bespoke business
rules that strip out known, erroneous data or provides a framework
against which to evaluate the likelihood of a fact being useful or
true. At the very least, there could be some work to develop ontologies
and schemas that detect inconsistencies in the data, but in a way
that adds value to a working system rather than such modelling being a
barrier to building the first iteration of the system.

Facts that are missing are arguably not a ``noise'' issue, but still
a data integrity concern. If an item lacks key information, it may
be categorised incorrectly. It should be noted that the data pipeline
designed here did nothing to address this deliberately and the desire
was to evaluate the open world characteristics of semantics without
trying to fill in the gaps.

As with the erroneous facts discussion above, there is much case to
be made for designing such a system so that it is extensible enough
with customisations and bespoke software to aid where it appears to be
struggling -- once its initial behaviour is evaluated. In the case
of the data pipeline designed in chapter~\ref{chp:design}, any
in-house databases and APIs can be integrated as additional extraction
techniques to be joined via union to more general approaches such
as dereferencing.

Data being irrelevant might be the larger of the three concerns here
as that greedy machine learning algorithms can get stuck on such
noise. In section~\ref{sec:few-large}, we see how one approach
caused the clustering to get stuck on the fact that all items linked
to a common FAQ page.

Other data attributes such as the Twitter
Cards metadata (see section~\ref{sec:smallest-cluster}) can contain
very broad things that are declared on all pages that use this
vocabulary (e.g. \texttt{card=summary} is always used to indicate that
a given page is using these meta tags) might cause a greey algorithm
to focus too long on grouping together all pages that happen to choose
to implement this Twitter feature with less concern about the more
interesting properties.
